Supplementary figures for Detailed disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis

### 2020-12-18 manuscript/manuscript.Rmd
Version sent to co-authors to reapprove for PLoS

The Covid-19 pandemic caused by severe acute respiratory syndrome coronavirus (SARS-CoV-2) has, as of November 2020, led to over 46 million cases and over 1.1 million deaths. Proposed treatments include antivirals approved for other indications (chloroquine, hydroxychloroquine, lopinavir/ritonavir, remdesivir, favipiravir, umifenovir), azithromycin, corticosteroids, immunoglobulins, tocilizumab and convalescent plasma [@http://zotero.org/users/5567156/items/8XWTUMUI; @http://zotero.org/users/5567156/items/6Q8Z8VJR]. As of this writing, the only treatment that exhibited positive effects in randomized trials is the corticosteroid dexamethasone for severe cases [@http://zotero.org/users/5567156/items/IW6VZ53Q; @http://zotero.org/users/5567156/items/5VAKEW58]. Remdesivir, previously designed and approved for Ebola, SARS and MERS, is currently the only antiviral against SARS-CoV-2 approved by the FDA. However, during spring and summer it was available only in limited amounts via compassionate use programme. Randomized clinical trials on remdesivir have provided  conflicting evidence [@http://zotero.org/users/5567156/items/MFS3IPPU; @http://zotero.org/users/5567156/items/KB5UDZH3]. Despite initial promise, hydroxychloroquine has failed to show benefit for hospitalized patients [@http://zotero.org/users/5567156/items/9X3UH8ZM; @http://zotero.org/users/5567156/items/KB5UDZH3], as an early treatment [@http://zotero.org/users/5567156/items/G6KZKVMI], as well as pre- and post-exposure prophylaxis [@http://zotero.org/users/5567156/items/7C23YCV5; @http://zotero.org/users/5567156/items/GMR9FX9Z; @http://zotero.org/users/5567156/items/H4CLJNUR]. Further, no improvement could be discerned when adding azithromycin to hydroxychloroquine therapy [@http://zotero.org/users/5567156/items/BQLRZGTQ]. Similarly, lopinavir/ritonavir did not show benefit [@http://zotero.org/users/5567156/items/6I6GAVWU]. No benefit was observed for convalescent plasma [@http://zotero.org/users/5567156/items/UEXLDJCB; @http://zotero.org/users/5567156/items/BHGT88US].

@@

```{r treatment_comb_table, echo=FALSE}
# Table \@ref(tab:treatment_comb_table} gives overview of combinations of treatments used in individual patients.
# Currently not showing - almost all of this information is visible in the table above. Will put in supplement
#pd<-data$patient_data
# x<-sort(table(paste0(
#   as.numeric(pd$ever_hcq),
#   as.numeric(pd$ever_az),
#   as.numeric(pd$ever_convalescent_plasma),
#   as.numeric(pd$ever_favipiravir),
#   as.numeric(pd$ever_remdesivir),
#   as.numeric(pd$ever_dexamethasone),
#   as.numeric(pd$ever_antibiotics))),dec=TRUE)
# x2<-c()
# nms<-c('HCQ','AZ','Plasma','Favipiravir','Remdesivir','Dexamethasone','ATB')
# for (i in 1:length(x)) {
#   x3<-c()
#   for (i2 in 1:nchar(names(x)[i])) {
#     x3<-c(x3,ifelse(substring(names(x)[i],i2,i2)=='1',nms[i2],''))
#   }
#   x2<-rbind(x2,cbind(n=x[i],rbind(x3)))
# }
# rownames(x2)<-c()
# colnames(x2)<-c('N',nms)
# x2<-x2[,c(1,2,3,8,4:7)]
# knitr::kable(x2,caption = "Treatment combinations in individual patients.", booktabs = TRUE)
```

@@

```{r progression, fig.cap="Disease progression for all patients included in the study as determined by breathing support required. Each horizontal strip is a single patient, the ordering on the vertical axis is by disease severity, the split into two panels is for visual purposes only. Ventilated = any form of ventilation (non-invasive positive-pressure ventilation, mechanical ventilation and extra-corporeal membrane oxygenation)."}
my_paste <- function(x) {
  paste0(as.integer(x), collapse = "")
}

@@

fig_progression <- data_for_fig_progression %>% 
  filter(!is_outcome) %>%
  ggplot(aes(x = day, y = patient_id, fill = breathing_s)) + 
  geom_tile(height = 0.7) + 
  geom_point(data = data_for_fig_progression %>% filter(is_outcome) %>% mutate(outcome = factor(breathing_s, ordered = FALSE)),
             inherit.aes = FALSE, aes(x = day, y = patient_id, shape = outcome), size = 0.9) +
  scale_fill_viridis_d("", direction = -1) +
  scale_x_continuous("Days since hospitalization") +
  # scale_color_manual("", values = c("black", "darkred"))+
  scale_shape_manual("", values = c(1, 15), guide = guide_legend(override.aes = list(size = 4))) +
  facet_wrap(~first_half, scales = "free_y") +
  theme(axis.text.y = element_blank(), strip.background = element_blank(), strip.text = element_blank(),
        legend.direction = "horizontal", legend.position = "bottom") +
  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank())

@@

As noted above, the nature of the convenience sample did not enforce random assignment of treatments to patients. In fact, patients with worse baseline characteristics, which lead to worse outcomes, were less likely to receive hydroxychloroquine (see Supplementary Figure 4). This clearly creates a bias towards a positive effect of hydroxychloroquine on the outcome (and potentially for other treatments as well - most were used in combination with hydroxychloroquine), which, however, could be false.

>>>>

The Covid-19 pandemic caused by severe acute respiratory syndrome coronavirus (SARS-CoV-2) has, as of November 2020, led to over 46 million cases and over 1.1 million deaths. Proposed treatments include antivirals approved for other indications (chloroquine, hydroxychloroquine, lopinavir/ritonavir, remdesivir, favipiravir, umifenovir), azithromycin, corticosteroids, immunoglobulins, tocilizumab and convalescent plasma [@http://zotero.org/users/5567156/items/8XWTUMUI; @http://zotero.org/users/5567156/items/6Q8Z8VJR]. 
As of this writing, the only treatment that exhibited positive effects in published randomized trials is the corticosteroid dexamethasone for severe cases [@http://zotero.org/users/5567156/items/IW6VZ53Q; @http://zotero.org/users/5567156/items/5VAKEW58], Janus kinase inhibitor baricitinib for hospitalized patients[@http://zotero.org/users/5567156/items/PDEPFSAU] and the monoclonal antibody bamlanivimab for non-hospitalized patients [@http://zotero.org/users/5567156/items/TZFUM6JH]. 
Other monoclonal antibodies have also announced, but not yet published positive results of randomized trials, including tocilizumab[@http://zotero.org/users/5567156/items/4SLQXUJM] and combination of casirivimab and imdevimab[@http://zotero.org/users/5567156/items/VTR54LI9].
Remdesivir, previously designed and approved for Ebola, SARS and MERS, is currently the only antiviral against SARS-CoV-2 approved by the FDA. However, during spring and summer it was available only in limited amounts via compassionate use programme. Randomized clinical trials on remdesivir have provided  conflicting evidence [@http://zotero.org/users/5567156/items/MFS3IPPU; @http://zotero.org/users/5567156/items/KB5UDZH3]. Despite initial promise, hydroxychloroquine has failed to show benefit for hospitalized patients [@http://zotero.org/users/5567156/items/9X3UH8ZM; @http://zotero.org/users/5567156/items/KB5UDZH3], as an early treatment [@http://zotero.org/users/5567156/items/G6KZKVMI], as well as pre- and post-exposure prophylaxis [@http://zotero.org/users/5567156/items/7C23YCV5; @http://zotero.org/users/5567156/items/GMR9FX9Z; @http://zotero.org/users/5567156/items/H4CLJNUR]. Further, no improvement could be discerned when adding azithromycin to hydroxychloroquine therapy [@http://zotero.org/users/5567156/items/BQLRZGTQ]. Similarly, lopinavir/ritonavir did not show benefit [@http://zotero.org/users/5567156/items/6I6GAVWU]. No benefit was observed for convalescent plasma [@http://zotero.org/users/5567156/items/UEXLDJCB; @http://zotero.org/users/5567156/items/BHGT88US].

@@

```{r}
median_ratios <-  data$marker_data %>% 
  filter(marker %in% c("lymphocyte_count", "d_dimer", "CRP", "IL_6")) %>%
  group_by(patient_id, marker) %>%
  filter(n() > 2) %>%
  summarise(ratio = max(value) / min(value), .groups = "drop") %>%
  group_by(marker) %>%
  summarise(med_ratio = median(ratio), .groups = "drop") %>%
  column_to_rownames("marker") %>% as.matrix() %>% round(1)
```

@@

The between-patient variability is notable even across outcomes - when ordering the patients by the highest CRP levels experienced throughout the hospital stay, the top 20% of patients that breathed ambient air for the whole hospitalization experienced higher levels than the bottom 20% of patients that required ventilation or died. This overlap is even larger when comparing only against the patients that died and D-dimer, Interleukin-6 and lymphocyte count also show a notably larger overlap than CRP (Supplementary Figure 4).

```{r progression, fig.cap="Disease progression for all patients included in the study as determined by breathing support required. Each vertical strip is a single patient, the ordering on the horizontal axis is by disease severity. Ventilated = any form of ventilation (non-invasive positive-pressure ventilation, mechanical ventilation and extra-corporeal membrane oxygenation)."}
my_paste <- function(x) {
  paste0(as.integer(x), collapse = "")
}

@@

fig_progression <- data_for_fig_progression %>% 
  mutate(patient_id = fct_reorder(patient_id, breathing_s,  .fun = my_paste, .desc = TRUE)) %>%
  filter(!is_outcome) %>%
  ggplot(aes(x = patient_id, y = day, fill = breathing_s)) + 
  geom_tile(width = 0.7) + 
  geom_point(data = data_for_fig_progression %>% filter(is_outcome) %>% mutate(outcome = factor(breathing_s, ordered = FALSE)),
             inherit.aes = FALSE, aes(x = patient_id, y = day, shape = outcome), size = 0.9) +
  scale_fill_viridis_d("", direction = -1) +
  scale_y_continuous("Days since hospitalization") +
  # scale_color_manual("", values = c("black", "darkred"))+
  scale_shape_manual("", values = c(1, 15), guide = guide_legend(override.aes = list(size = 4))) +
  theme(axis.text.x = element_blank(), strip.background = element_blank(), strip.text = element_blank(),
        legend.direction = "horizontal", legend.position = "bottom") +
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank())

@@

As noted above, the nature of the convenience sample did not enforce random assignment of treatments to patients. In fact, patients with worse baseline characteristics, which lead to worse outcomes, were less likely to receive hydroxychloroquine (see Figure&nbsp;\@ref(fig:hcqassoc)). This clearly creates a bias towards a positive effect of hydroxychloroquine on the outcome (and potentially for other treatments as well - most were used in combination with hydroxychloroquine), which, however, could be false.

```{r hcqassoc, message=FALSE, fig.width=7, fig.height=4.5, fig.cap="The choice of treatment with hydroxychloroquine seemed to be associated with the status of patients upon hospitalization. Comorbidities were associated with both worse outcome (black) and lower chance of treatment with hydroxychloroquine (red). Dots and lines represent the estimates and the 95% confidence intervals of the log odds ratio of the respective outcome. HCQ: hydroxychloroquine, IHD: ischemic heart disease, HD: hypertension drugs, HF: heart failure history, COPD: chronic obstructive pulmonary disease, LungD: other lung disease, Dia: diabetes, RenalD: renal disease, LiverD: liver disease, HighCr: creatinin above 115 for males or above 97 for females, HighInr: Prothrombin time (Quick test) as International Normalized Ratio above 1.2, LowAlb: albumin in serum/plasma below 36 g/l."}

tmp<-data$patient_data[,c('age','sex','days_from_symptom_onset','admitted_for_covid','ischemic_heart_disease','has_hypertension_drugs','heart_failure',
  'COPD','asthma','other_lung_disease','diabetes','renal_disease','liver_disease','smoking','BMI','obesity','last_record','high_creatinin','high_pt_inr','low_albumin',
  'heart_problems','outcome','ever_hcq','ever_az','worst_breathing','had_invasive'),]
tmp$outcome[tmp$outcome=='Transferred']<-'Hospitalized'
tmp$outcome<-dropLevels(tmp$outcome)
tmp2<-tmp[tmp$outcome%in%c('Death','Discharged'),]
tmp2$outcome<-dropLevels(tmp2$outcome)
tmp2$outcome<-tmp2$outcome=='Death'

m.ihd<-glm(outcome~ischemic_heart_disease,tmp2,family='binomial')
m.hd<-glm(outcome~has_hypertension_drugs,tmp2,family='binomial')
m.hf<-glm(outcome~heart_failure,tmp2,family='binomial')
m.copd<-glm(outcome~COPD,tmp2,family='binomial')
m.asthma<-glm(outcome~asthma,tmp2,family='binomial')
m.old<-glm(outcome~other_lung_disease,tmp2,family='binomial')
m.dia<-glm(outcome~diabetes,tmp2,family='binomial')
m.rd<-glm(outcome~renal_disease,tmp2,family='binomial')
m.ld<-glm(outcome~liver_disease,tmp2,family='binomial')
m.s<-glm(outcome~smoking,tmp2,family='binomial')
m.ob<-glm(outcome~obesity,tmp2,family='binomial')
m.hc<-glm(outcome~high_creatinin,tmp2,family='binomial')
m.hi<-glm(outcome~high_pt_inr,tmp2,family='binomial')
m.la<-glm(outcome~low_albumin,tmp2,family='binomial')

m2.ihd<-glm(ever_hcq~ischemic_heart_disease,tmp2,family='binomial')
m2.hd<-glm(ever_hcq~has_hypertension_drugs,tmp2,family='binomial')
m2.hf<-glm(ever_hcq~heart_failure,tmp2,family='binomial')
m2.copd<-glm(ever_hcq~COPD,tmp2,family='binomial')
m2.asthma<-glm(ever_hcq~asthma,tmp2,family='binomial')
m2.old<-glm(ever_hcq~other_lung_disease,tmp2,family='binomial')
m2.dia<-glm(ever_hcq~diabetes,tmp2,family='binomial')
m2.rd<-glm(ever_hcq~renal_disease,tmp2,family='binomial')
m2.ld<-glm(ever_hcq~liver_disease,tmp2,family='binomial')
m2.s<-glm(ever_hcq~smoking,tmp2,family='binomial')
m2.ob<-glm(outcome~obesity,tmp2,family='binomial')
m2.hc<-glm(ever_hcq~high_creatinin,tmp2,family='binomial')
m2.hi<-glm(ever_hcq~high_pt_inr,tmp2,family='binomial')
m2.la<-glm(ever_hcq~low_albumin,tmp2,family='binomial')

models<-list(m.ihd,m.hd,m.hf,m.copd,m.asthma,m.old,m.dia,m.rd,m.ld,m.s,m.ob,m.hc,m.hi,m.la)
models2<-list(m2.ihd,m2.hd,m2.hf,m2.copd,m2.asthma,m2.old,m2.dia,m2.rd,m2.ld,m2.s,m2.ob,m2.hc,m2.hi,m2.la)
modelNames<-c('IHD','HD','HF','COPD','Asthma','LungD','Dia','RenalD','LiverD','Smoking','Obesity','HighCr','HighInr','LowAlb')
gtmp<-c()
for (i in rev(seq(along=models))) {
  ii<-length(models)-i+1
  m<-models[[i]]
  m2<-models2[[i]]
  ci<-confint(m)[2,]
  ci2<-confint(m2)[2,]
  gtmp<-rbind(gtmp,data.frame(est=coef(m)[2],low=ci[1],high=ci[2],name=modelNames[i],type='outcome'))
  gtmp<-rbind(gtmp,data.frame(est=-coef(m2)[2],low=-ci2[1],high=-ci2[2],name=modelNames[i],type='hcq'))
}
ggplot(gtmp,aes(y=name,x=est,xmin=low,xmax=high,color=type))+
  geom_vline(xintercept=0,color='blue')+
  geom_point(position=position_dodge(width=0.5))+
  geom_errorbarh(position=position_dodge(width=0.5),height= 0.5)+
  scale_x_continuous('Log odds ratio')+
  scale_y_discrete('')+
  scale_color_manual('Model of:',values=c('black','red'),labels=c('Death','No HCQ'))+
  theme(legend.background = element_rect(fill="gray95"),legend.position=c(.83,.7))
```

### 2020-12-18 manuscript/supplementary_figures.Rmd
Tweaks VK, formatting fixes for the supplements
Version sent to co-authors to reapprove for PLoS

```{r upsettreatment, fig.cap = "Upset plot of treatment combinations - each vertical bar displays the number of patients that received the combination indicated by filled dots in the matrix. Horizontal bars show the total number of patients receiveing the given treatment.", fig.width=8, fig.height=6 }
upset_data <- data$patient_data %>% select(starts_with("ever")) %>% mutate(across(starts_with("ever"), as.integer)) %>% as.data.frame()

@@

```{r hospitaloutcomes,fig.cap="Number of patients and outcomes at the individual sites. The numbers above bars are the exact counts. Hospitalized = still hospitalized at the end of data collection at the site. Sites are anonymized to preserve patient privacy."}
data$patient_data %>% mutate(
  hospital_label = factor(paste0("H", as.integer(factor(hospital_id))), levels = paste0("H", 1:10)),
  outcome = fct_collapse(outcome, Hospitalized = c("Hospitalized", "Transferred")) %>% fct_relevel("Death", "Hospitalized", "Discharged")
) %>%
  ggplot(aes(x = hospital_label,  fill = outcome)) + geom_bar(position = position_dodge()) + 
  geom_text(stat='count', aes(label=..count..), vjust=-1, position = position_dodge(width = 1)) +
  #scale_fill_viridis_d("Outcome") +
  scale_fill_manual("Outcome", values = c(Death = as.character(bfc()["Black"]), Hospitalized = as.character(bfc()["Bluish green"]), Discharged = as.character(bfc()["Yellow"]))) +
  scale_x_discrete("Study site") +
  scale_y_continuous("Count") +
  expand_limits(y = 36)
```

```{r hcqassoc, message=FALSE, fig.width=7, fig.height=4.5, fig.cap="The choice of treatment with hydroxychloroquine seemed to be associated with the status of patients upon hospitalization. Comorbidities were associated with both worse outcome (black) and lower chance of treatment with hydroxychloroquine (red). Dots and lines represent the estimates and the 95% confidence intervals of the log odds ratio of the respective outcome. HCQ: hydroxychloroquine, IHD: ischemic heart disease, HD: hypertension drugs, HF: heart failure history, COPD: chronic obstructive pulmonary disease, LungD: other lung disease, Dia: diabetes, RenalD: renal disease, LiverD: liver disease, HighCr: creatinin above 115 for males or above 97 for females, HighInr: Prothrombin time (Quick test) as International Normalized Ratio above 1.2, LowAlb: albumin in serum/plasma below 36 g/l."}

tmp<-data$patient_data[,c('age','sex','days_from_symptom_onset','admitted_for_covid','ischemic_heart_disease','has_hypertension_drugs','heart_failure',
  'COPD','asthma','other_lung_disease','diabetes','renal_disease','liver_disease','smoking','BMI','obesity','last_record','high_creatinin','high_pt_inr','low_albumin',
  'heart_problems','outcome','ever_hcq','ever_az','worst_breathing','had_invasive'),]
tmp$outcome[tmp$outcome=='Transferred']<-'Hospitalized'
tmp$outcome<-dropLevels(tmp$outcome)
tmp2<-tmp[tmp$outcome%in%c('Death','Discharged'),]
tmp2$outcome<-dropLevels(tmp2$outcome)
tmp2$outcome<-tmp2$outcome=='Death'

m.ihd<-glm(outcome~ischemic_heart_disease,tmp2,family='binomial')
m.hd<-glm(outcome~has_hypertension_drugs,tmp2,family='binomial')
m.hf<-glm(outcome~heart_failure,tmp2,family='binomial')
m.copd<-glm(outcome~COPD,tmp2,family='binomial')
m.asthma<-glm(outcome~asthma,tmp2,family='binomial')
m.old<-glm(outcome~other_lung_disease,tmp2,family='binomial')
m.dia<-glm(outcome~diabetes,tmp2,family='binomial')
m.rd<-glm(outcome~renal_disease,tmp2,family='binomial')
m.ld<-glm(outcome~liver_disease,tmp2,family='binomial')
m.s<-glm(outcome~smoking,tmp2,family='binomial')
m.ob<-glm(outcome~obesity,tmp2,family='binomial')
m.hc<-glm(outcome~high_creatinin,tmp2,family='binomial')
m.hi<-glm(outcome~high_pt_inr,tmp2,family='binomial')
m.la<-glm(outcome~low_albumin,tmp2,family='binomial')

m2.ihd<-glm(ever_hcq~ischemic_heart_disease,tmp2,family='binomial')
m2.hd<-glm(ever_hcq~has_hypertension_drugs,tmp2,family='binomial')
m2.hf<-glm(ever_hcq~heart_failure,tmp2,family='binomial')
m2.copd<-glm(ever_hcq~COPD,tmp2,family='binomial')
m2.asthma<-glm(ever_hcq~asthma,tmp2,family='binomial')
m2.old<-glm(ever_hcq~other_lung_disease,tmp2,family='binomial')
m2.dia<-glm(ever_hcq~diabetes,tmp2,family='binomial')
m2.rd<-glm(ever_hcq~renal_disease,tmp2,family='binomial')
m2.ld<-glm(ever_hcq~liver_disease,tmp2,family='binomial')
m2.s<-glm(ever_hcq~smoking,tmp2,family='binomial')
m2.ob<-glm(outcome~obesity,tmp2,family='binomial')
m2.hc<-glm(ever_hcq~high_creatinin,tmp2,family='binomial')
m2.hi<-glm(ever_hcq~high_pt_inr,tmp2,family='binomial')
m2.la<-glm(ever_hcq~low_albumin,tmp2,family='binomial')

models<-list(m.ihd,m.hd,m.hf,m.copd,m.asthma,m.old,m.dia,m.rd,m.ld,m.s,m.ob,m.hc,m.hi,m.la)
models2<-list(m2.ihd,m2.hd,m2.hf,m2.copd,m2.asthma,m2.old,m2.dia,m2.rd,m2.ld,m2.s,m2.ob,m2.hc,m2.hi,m2.la)
modelNames<-c('IHD','HD','HF','COPD','Asthma','LungD','Dia','RenalD','LiverD','Smoking','Obesity','HighCr','HighInr','LowAlb')
gtmp<-c()
for (i in rev(seq(along=models))) {
  ii<-length(models)-i+1
  m<-models[[i]]
  m2<-models2[[i]]
  ci<-confint(m)[2,]
  ci2<-confint(m2)[2,]
  gtmp<-rbind(gtmp,data.frame(est=coef(m)[2],low=ci[1],high=ci[2],name=modelNames[i],type='outcome'))
  gtmp<-rbind(gtmp,data.frame(est=-coef(m2)[2],low=-ci2[1],high=-ci2[2],name=modelNames[i],type='hcq'))
}
ggplot(gtmp,aes(y=name,x=est,xmin=low,xmax=high,color=type))+
  geom_vline(xintercept=0,color='blue')+
  geom_point(position=position_dodge(width=0.5))+
  geom_errorbarh(position=position_dodge(width=0.5),height= 0.5)+
  scale_x_continuous('Log odds ratio')+
  scale_y_discrete('')+
  scale_color_manual('Model of:',values=c('black','red'),labels=c('Death','No HCQ'))+
  theme(legend.background = element_rect(fill="gray95"),legend.position=c(.83,.7))
```

>>>>

```{r upsettreatment, fig.cap = "Upset plot of treatment combinations - each vertical bar displays the number of patients that received the combination indicated by filled dots in the matrix. Horizontal bars show the total number of patients receiveing the given treatment.", fig.width=8, fig.height=5.8 }
upset_data <- data$patient_data %>% select(starts_with("ever")) %>% mutate(across(starts_with("ever"), as.integer)) %>% as.data.frame()

@@

```{r hospitaloutcomes,fig.cap="Number of patients and outcomes at the individual sites. The numbers above bars are the exact counts. Hospitalized = still hospitalized at the end of data collection at the site or transferred to other site and lost to followup. Sites are anonymized to preserve patient privacy.", fig.height=3.5}
data$patient_data %>% mutate(
  hospital_label = factor(paste0("H", as.integer(factor(hospital_id))), levels = paste0("H", 1:10)),
  outcome = fct_collapse(outcome, Hospitalized = c("Hospitalized", "Transferred")) %>% fct_relevel("Death", "Hospitalized", "Discharged")
) %>%
  ggplot(aes(x = hospital_label,  fill = outcome)) + geom_bar(position = position_dodge()) + 
  geom_text(stat='count', aes(label=..count..), vjust=-1, position = position_dodge(width = 1)) +
  #scale_fill_viridis_d("Outcome") +
  scale_fill_manual("Outcome", values = c(Death = as.character(bfc()["Black"]), Hospitalized = as.character(bfc()["Bluish green"]), Discharged = as.character(bfc()["Yellow"]))) +
  scale_x_discrete("Study site") +
  scale_y_continuous("Count") +
  expand_limits(y = 36)
```

```{r markeroverlaps, fig.cap = "Density plots of worst marker values per patient, stratified by worst condition experienced by the patient.", fig.height=6}
frac <- 0.2
data$marker_data %>% filter(marker %in% c("lymphocyte_count", "d_dimer", "CRP", "IL_6")) %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>%
  #filter(worst_breathing %in% c("AA", "Death")) %>%
  group_by(marker, patient_id, worst_breathing_s) %>%
  summarise(worst_value = case_when(marker == "lymphocyte_count" ~ min(log10(value)),
                                    marker == "IL_6" ~ max(log10(value)),
                                    TRUE ~ max(value)),
            .groups = "drop") %>% 
  filter(!is.infinite(worst_value)) %>%
  mutate(marker = factor(marker, levels = c("lymphocyte_count", "d_dimer", "CRP", "IL_6"),
         labels = c("Lowest Log10(lymphocyte count)", "Highest D-dimer", "Highest CRP", "Highest Log10(IL_6)"))) %>%
  ggplot(aes(x = worst_value, color = worst_breathing_s, fill = worst_breathing_s, group = worst_breathing_s)) +
  geom_density(adjust = 2, alpha = 0.3) +
  scale_color_viridis_d("Worst condition") +
  scale_fill_viridis_d("Worst condition") +
  scale_x_continuous("Marker value") +
  facet_wrap(~marker, scales = "free", ncol = 1)

#Validating the claim about overlapping proportions
# frac <- 0.2
# data$marker_data %>% filter(marker %in% c("lymphocyte_count", "d_dimer", "CRP")) %>%
#   inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>%
#   filter(worst_breathing_s != "Oxygen") %>%
#   mutate(worst_breathing_s = fct_collapse(worst_breathing_s, "Ventilated or Death" = c("Ventilated", "Death"))) %>%
#   group_by(marker, patient_id, worst_breathing_s) %>%
#   summarise(min_value = min(value), max_value = max(value), .groups = "drop") %>%
#   group_by(marker, worst_breathing_s) %>%
#   summarise(qlowmin = quantile(min_value, frac), qhigh_min = quantile(min_value, 1 - frac), qlow_max = quantile(max_value, frac), qhigh_max = quantile(max_value, 1 - frac), .groups = "drop")

```
```{r}
# data$marker_data %>% 
#   filter(marker %in% c("lymphocyte_count", "CRP", "d_dimer","IL_6"), value > 0, day >=0) %>%
#   inner_join(data$breathing_data, by = c("hospital_id", "patient_id", "day")) %>%
#   filter(breathing_s != "Death") %>%
#   ggplot(aes(x = day, y = value, color = breathing_s, group = patient_id)) +
#     geom_line(alpha = 0.8, position = position_jitter(width = 0.2, height = 0)) +
#    # geom_point(, alpha = 0.5, position = position_jitter(width = 0.2, height = 0)) +
#   scale_y_log10() + 
#   scale_color_manual("Breathing", values = c(Ventilated = as.character(bfc()["Black"]), Oxygen = as.character(bfc()["Bluish green"]), AA = as.character(bfc()["Orange"]))) + #scale_color_brewer(type = "qual") +
#   # scale_color_viridis_d(guide = guide_legend(override.aes = list(alpha = 1)), direction = -1, end = 0.9) +
#   facet_wrap(~marker, scales = "free_y")
#   
```
